Implementation of Hadoop 2 on PBS-based HPC Systems

### 2014-04-15 paper/Paper.tex
Added refrences from Alex.
Fixed spelling errors.

	\begin{abstract}
		Mark Baker, Alex Berk, and Joshua Hull implemented Hadoop 2 on a PBS-based HPC System on the Palmetto Cluster in a version update from Hadoop 1. Development saw challenges to creating a cluster environment on the Palmetto cluster that maintained compatibility with existing Hadoop 1 scripts. Infrastructure and PBS script editing were necessary to support the additional features and YARN components in the Hadoop 2 framework. Thorough testing with WordCount and TeraSort was used to measure the success of implementation and initialization of the Hadoop 2 cluster.
	\end{abstract}

	\section{Introduction}
		As part of a group project for Clemson University's class on Distributed and Cluster Computing Mark Baker, Alex Berk, and Joshua Hull worked to implement the myHadoop2 framework and environment on the Palmetto Supercomputing Cluster at Clemson University. This environment will allow users to run the updated Hadoop 2 framework in an on demand cluster on Palmetto. Hadoop 2 introduced additional tools that need to be handled by myHadoop2, such as the YARN resource management system. These new tools help to reduce bottlenecks in Hadoop and generally increase performance. The purpose of myHadoop2 is to give the users of Palmetto and other PBS based systems a way to set up a dynamic Hadoop 2 cluster in order to take advantage of these new features while still allowing use of the original Hadoop 1 programs.

@@

			Alex was primarily responsible for the development of the myHadoop2 script with assistance from Joshua. Alex is currently working with Clemson Computing and Information Technology, or CCIT, to get Hadoop 2 working on the Palmetto cluster, and was able to bring this experience to the project and successfully implement that goal for both the success of the project and the goal of CCIT. He is directly involved in the implementation of Hadoop 2 over OrangeFS as well, which greatly benefited him in progressing the myHadoop scripting on the cluster efficiently and with the best performance.

			Mark was responsible for documentation of the project and collaboration on development with Alex and Joshua with the implementation to complete this objective. Mark has worked as a functional analyst for the past three years and has experience in documentation of analyses and projects. He worked closely with Joshua and Alex throughout the development and implementation processes to bring a complete and detailed understanding of the project to its documentation and presentation.

@@

                With the new features prevalent in Hadoop 2, there was also the issue of implementing the new resource tracking system that wasn’t initially set up when creating the system in our project, but with the work Alex had been working on for CCIT to set up their own configuration of Hadoop 2, we were able to borrow the OpenClemson source\cite{openclemson} for the resource system to assist us in getting ours working correctly.
			\subsection{Palmetto Environment}
				One of our earliest challenges was in setting up the new Hadoop environment on the Palmetto without disrupting the current Hadoop 1 framework and optimizing Hadoop 2 to run its configuration in compatibility with the current Hadoop. Fortunately, Alex had been working with CCIT in Clemson on getting this done with OrangeFS. Through research and contact, he found a partially configured open source implementation of Hadoop 2.2.0 developed by the San Diego Supercomputer Center. With a few bug fixes to let certain conditions on PBS Pro work correctly, we managed to incorporate the new myHadoop in our scripting and testing without having to greatly edit or modify an older version. The initial release of myHadoop 0.20 provided configuration files that edited the original Hadoop files for compatibility with PBS systems. This approach was difficult to read and understand, requiring the programmer to investigate each specific file for changes. The new myHadoop 0.30 has been completely reworked for simplicity and scalability.  A patch file is used to alter the Hadoop configuration depending on Hadoop version, and a universal configuration file is used to generate the appropriate Hadoop xml configuration files. As a result, we now have a well-organized environment for Hadoop users.

@@

			\subsection{Hadoop 2 Testing}
               			\subsubsection{WordCount}
                    				The first benchmark conducted was WordCount. This benchmark counted the number of words in several classic works from the Project Gutenberg website\cite{gutenburg}:
                    				\begin{itemize}
                    				 	\item A Connecticut Yankee in King Arthur's Court\cite{yankee}
                        					\item Adventures of Huckleberry Finn\cite{huck}
                        					\item Around The World in 80 Days\cite{80days}
                      					\item From The Earth to the Moon\cite{moon}
                     					\item The Adventures of Tom Sawyer\cite{tom}
                  				\end{itemize}
						The benchmark used the unmodified WordCount example code from the Hadoop 1.1.2 downloaded from Apache and compiled into a java jar file. The jar file was run via the hadoop command and timed via the time command. The recorded user times are plotted in figure one below.

		Each compute node used 17 processing cores and 10 GB of memory and was networked to the other nodes via a 10 GB Ethernet link. One node was designated the headnode and ran the NameNode, JobTracker, and YARN resource manager components of Hadoop. Each node, including the headnode, also ran the TaskTracker and DataNode components.

@@

					As previously seen with the WordCount tests, TeraSort showed increased runtimes with increased nodes due to overheads early but improved after several TaskTracker nodes were added. Furthermore, we see a significant decrease from a sequential, singular node running TeraSort in comparison with two nodes, but a sharp increase when moving up to eight nodes. This could be caused by many factors such as resource management not being optimal during testing or communications being congested since the Palmetto is always open for other users to run their programs. This could cause small, but consistent, performance drops in our runtimes.
				\subsection{Hadoop 1 Testing}
					The Hadoop benchmarking took place under similar conditions to the Hadoop 2 tests. There was one node specified at the headnode that ran the NameNode and JobTracker components to Hadoop. The primary difference is that the headnode did not run TaskTracker or DataNode components. Each node used 17 processor cores and 10 GB of memory. The same java jar file was used to benchmark Hadoop as was used in Hadoop 2. The runtimes were collected in a similar manner to Hadoop 2.
					\subsubsection{WordCount}
					WordCount on Hadoop was run using the same set of input files as listed above with Hadoop 2. The only changes were those needed due to Hadoop version differences. These include using hadoop commands that were deprecated with Hadoop 2.

@@

					We can see that moving from sequential to multiple nodes improved runtime efficiency on Hadoop 1 as well and even had similar effects in overhead resource slowing down performance in the smaller number of node tests that improved later on. However, the Hadoop 1 tests could not replicate the testing circumstances of Hadoop 2 for larger amounts of datanodes due to Hadoop 1 and the Palmetto architecture working together. Because Hadoop 1 has a namenode setup that does not allow it to be a data processing node as well, in order to achieve a 16 node performance the system must allocate 17 nodes instead to account for the non-working namenode. In Palmetto’s current architecture, accessing more than 16 nodes requires allocation to a specific set of nodes on the cluster. These resources are much harder to acquire because of the uncommon request compared to many of the uses Palmetto sees. Therefore testing in Hadoop 1 to optimize time and efficiency could only test up to 8 working datanodes and so performance of a higher amount of nodes must be inferred based on
available data.
					\subsubsection{TeraSort}
						The TeraSort benchmarks were performed in a similar matter the Hadoop 2 Benchmarks. TeraGen was given 100000 as an input parameter.
						\begin{figure}[h]
                   					\begin{center}
                       						\input{hadoop_terasort}
                        						\caption{Hadoop TeraSort Runtimes}
                    				\end{center}
                  			\end{figure}

>>>>

	\begin{abstract}
		Mark Baker, Alex Berk, and Joshua Hull implemented Hadoop 2 on a PBS-based HPC System on the Palmetto Cluster in a version update from Hadoop 1. Development saw challenges to creating a cluster environment on the Palmetto cluster that maintained compatibility with existing Hadoop 1 programs. Infrastructure and PBS script editing were necessary to support the additional features and YARN components in the Hadoop 2 framework. Thorough testing with WordCount and TeraSort was used to measure the success of implementation and initialization of the Hadoop 2 cluster.
	\end{abstract}

	\section{Introduction}
		As part of a group project for Clemson University's class on Distributed and Cluster Computing Mark Baker, Alex Berk, and Joshua Hull worked to implement the myHadoop2 framework and environment on the Palmetto Supercomputing Cluster at Clemson University. This environment will allow users to run the updated Hadoop 2 framework in an on demand cluster on Palmetto. Hadoop 2 introduced additional tools that need to be handled by myHadoop2, such as the YARN resource management system. These new tools help to reduce bottlenecks in Hadoop and generally increase performance. The purpose of myHadoop2 is to give the users of Palmetto and other PBS based systems a way to set up a dynamic Hadoop 2 cluster in order to take advantage of these new features while still allowing use of Hadoop 1 programs.

@@

			Alex was primarily responsible for the development of the myHadoop2 script with assistance from Joshua. Alex is currently working with Clemson Computing and Information Technology, or CCIT, to get Hadoop 2 working on the Palmetto cluster, and was able to bring this experience to the project and successfully implement that goal for both the success of the project and the goal of CCIT. He is directly involved in the implementation of Hadoop 2 over OrangeFS as well, which greatly benefited him in progressing the myHadoop scripting on the cluster efficiently and with the best performance. He spent much of the Spring 2014 academic semester researching implementations and configurations for old and new Hadoop versions\cite{hadoopguide}\cite{hadoopyarn}

			Mark was responsible for documentation of the project and collaboration on development with Alex and Joshua with the implementation to complete the project's objective. Mark has worked as a functional analyst for the past three years and has experience in documentation of analyses and projects. He worked closely with Joshua and Alex throughout the development and implementation processes to bring a complete and detailed understanding of the project to its documentation and presentation.

@@

                With the new features prevalent in Hadoop 2, there was also the issue of implementing the new resource tracking system that wasn’t initially set up when creating the system in our project, but with the work Alex had been working on for CCIT to set up their own configuration of Hadoop 2, we were able to borrow the OpenClemson source\cite{openclemson} for the resource system to assist us in getting ours working correctly.
			\subsection{Palmetto Environment}
				One of our earliest challenges was in setting up the new Hadoop environment on the Palmetto without disrupting the current Hadoop 1 framework and optimizing Hadoop 2. Fortunately, Alex had been working with CCIT in Clemson on getting this done with OrangeFS. Through research and contact, he found a partially configured open source implementation of myHadoop that was designed to work with Hadoop 2.2.0 developed by the San Diego Supercomputer Center. With a few bug fixes to let certain conditions on PBS Pro work correctly, we managed to incorporate the new myHadoop in our scripting and testing without having to greatly edit or modify an older version. The initial release of myHadoop 0.20 provided configuration files that edited the original Hadoop files for compatibility with PBS systems. This approach was difficult to read and understand, requiring the programmer to investigate each specific file for changes. The new myHadoop 0.30 has been completely reworked for simplicity and scalability.  A patch file is used to alter the Hadoop configuration depending on Hadoop version, and a universal configuration file is used to generate the appropriate Hadoop xml configuration files. As a result, we now have a well-organized environment for Hadoop users.

@@

		The hardware configuration for each compute node was kept consistent during the testing. Each compute node used 17 processing cores and 10 GB of memory and was networked to the other nodes via a 10 GB Ethernet link. One node was designated the headnode and ran the NameNode, JobTracker, and YARN resource manager components of Hadoop. Each node, including the headnode, also ran the TaskTracker and DataNode components.
			\subsection{Hadoop 2 Testing}
               			\subsubsection{WordCount}
                    				The first benchmark conducted was WordCount. This benchmark counted the number of words in several classic works from the Project Gutenberg website\cite{gutenburg}:
                    				\begin{itemize}
                    				 	\item A Connecticut Yankee in King Arthur's Court\cite{yankee}
                        					\item Adventures of Huckleberry Finn\cite{huck}
                        					\item Around The World in 80 Days\cite{80days}
                      					\item From The Earth to the Moon\cite{moon}
                     					\item The Adventures of Tom Sawyer\cite{tom}
                  				\end{itemize}
						The benchmark used the unmodified WordCount example code from the Hadoop 1.1.2 downloaded from Apache and compiled into a Java jar file. The jar file was run via the hadoop command and timed via the time command. The recorded user times are plotted in figure one below.

@@

					As previously seen with the WordCount tests, TeraSort showed increased runtimes with increased nodes due to overheads early but improved after several TaskTracker nodes were added. Furthermore, we see a significant decrease from a sequential, singular node running TeraSort in comparison with two nodes, but a sharp increase when moving up to eight nodes. This could be caused by many factors such as resource management not being optimal during testing or communications being congested since the Palmetto is always open for other users to run their programs. This could cause small, but consistent, performance drops in our runtimes.
				\subsection{Hadoop 1 Testing}
					The Hadoop benchmarking took place under similar conditions to the Hadoop 2 tests. There was one node specified at the headnode that ran the NameNode and JobTracker components to Hadoop. The primary difference is that the headnode did not run the TaskTracker or DataNode components. Each node used 17 processor cores and 10 GB of memory. The same Java jar file was used to benchmark Hadoop as was used in Hadoop 2. The runtimes were collected in a similar manner to Hadoop 2.
					\subsubsection{WordCount}
					WordCount on Hadoop was run using the same set of input files as listed above with Hadoop 2. The only changes were those needed due to Hadoop version differences. These include using hadoop commands that were deprecated with Hadoop 2.

@@

					We can see in figure three that moving from sequential to multiple nodes improved runtime efficiency on Hadoop 1 as well and even had similar effects in overhead resource slowing down performance in the smaller number of node tests that improved later on. However, the Hadoop 1 tests could not replicate the testing circumstances of Hadoop 2 for larger amounts of datanodes due to Hadoop 1 and the Palmetto architecture working together. Because Hadoop 1 has a namenode setup that does not allow it to be a data processing node as well, in order to achieve a 16 node performance the system must allocate 17 nodes instead to account for the non-working namenode. In Palmetto’s current architecture, accessing more than 16 nodes requires allocation to a specific set of nodes on the cluster. These resources are much harder to acquire because of the uncommon request compared to many of the uses Palmetto sees. Therefore testing in Hadoop 1 to optimize time and efficiency could only test up to 8 working datanodes and so performance of a higher amount of nodes must be inferred based on
available data.
					\subsubsection{TeraSort}
						The TeraSort benchmarks were performed in a similar matter the Hadoop 2 Benchmarks. TeraGen was given 100000 as an input parameter.
						\begin{figure}[h]
                   					\begin{center}
                       						\input{hadoop_terasort}
                        						\caption{Hadoop TeraSort Runtimes}
                    				\end{center}
                  			\end{figure}
